Draft_figures_2020-06-02

### 2020-06-05 R/Draft_figures_2020-06-02.Rmd
4WCR F0 F1 and F2 figures and comparisons

J045.A SOURCE_ID_2 NSL 29896, Summer
```{r}
phe_F0_up <- phe_all %>%
  filter(PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
phe_F0_low <- phe_all %>%
  filter(PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
phe_F0 <- phe_all %>%
  filter(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
phe_F1 <- phe_all %>%
  filter(PLANT_ID %in% c("AxD", "VxW")) 
phe_F2 <- phe_all %>%
  filter(!(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
```

@@

>>>>

J045.A SOURCE_ID_2 NSL 29896, Summer
```{r}
phe_F0_up <- phe_all %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("DAC", "VS16", "J440.A", "J036.A", "J037.A")) 
phe_F0_low <- phe_all %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "WBC", "J230.A")) 
phe_F0 <- phe_all %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "J440.A", "J440.B", "J230.A", "J292.A", "J045.A", "J036.A", "J037.A", "J022.B", "J210.A")) 
phe_F1 <- phe_all %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AxD", "VxW")) 
phe_F2 <- phe_all %>%
  filter(WHERE == "FWCR" & !(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW"))) 
```

@@

# --------------
# 4WCR Plots

We then planted the four parents, the two F1 individuals (AP13xDAC, and VS16xWBC), and XXX F2 individuals at eight field sites, and recorded greenup and flowering date for the 2016-2019 seasons. Though there was a X day gap in flowering date between the two F1 crosses, F1 individuals flowered at similar dates as the Midwest parents and Midwest subpopulation individuals in general, indicating that the flowering time genetic response to the Midwestern GDD levels was dominant to the photoperiod response and the higher GDD levels required by the Gulf subpopulation. 
```{r}
phe_F1 %>%
  filter(PHE == "FL50" & YEAR == 2019) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(fl50 = mean(MEAS, na.rm = TRUE)) %>%
  pivot_wider(names_from = PLANT_ID, values_from = fl50) %>%
  mutate(diff = VxW - AxD) %>%
  group_by(SITE) %>%
  summarise(meandiff = mean(diff)) %>%
  arrange(meandiff) %>%
  filter(!(SITE %in% c("OVTN", "MNHT"))) %>%
  summarise(meandiff = mean(meandiff))

F1plot <- phe_F1 %>%
  filter(PHE == "FL50" & (YEAR %in% c(2019) | (SITE == "TMPL" & YEAR == 2018)) & !(SITE %in% c("OVTN", "MNHT"))) %>%
  left_join(sites, by = "SITE") %>%
  group_by(manu_site, YEAR, PLANT_ID) %>%
  summarise(MEAS = mean(MEAS, na.rm = TRUE))
phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = rev(c("4X", "Atlantic", "Gulf", "Midwest", "8X")))
color_code <- c("grey", "#6E91CB","#F47F72", "#442C83", "#090906")

F1vsGWAS <- phe_all %>%
  left_join(sites, by = "SITE") %>%
  filter(PHE == "FL50" & WHERE == "GWAS" & !(SITE %in% c("FRMI", "OVTN")) & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = as_date(MEAS, origin = "2019-01-01"))) +
  geom_vline(xintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_density(aes(color = SUBPOP)) + 
  facet_wrap(~manu_site, nrow = 2) +
  geom_vline(data = F1plot, aes(xintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  scale_color_manual(values = c("grey", "#F47F72", "#442C83",  "#090906")) +
  theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  ylab("Subpop Fraction") + xlab("Flowering Date") +
  geom_hline(yintercept = 0, color = "darkgrey")
save_plot(filename = "../manuscript/plots/F1_vs_GWAS_distribution.png", plot = F1vsGWAS, base_asp = 2.5, base_height = 2.5)

F1vsGWAS_violin <- phe_all %>%
  left_join(sites, by = "SITE") %>%
  filter(PHE == "FL50" & WHERE == "GWAS" & !(SITE %in% c("FRMI", "OVTN")) & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = SUBPOP, y = as_date(MEAS, origin = "2019-01-01"))) +
  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_violin(aes(fill = SUBPOP)) + 
  facet_wrap(~manu_site, nrow = 2) +
  geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  scale_color_manual(values = c("grey",  "#090906", "#F47F72", "#442C83")) +
  scale_fill_manual(values = c("#442C83", "#F47F72")) +
  theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  xlab("Subpopulation") + ylab("Flowering Date") + coord_flip()

save_plot(filename = "../manuscript/plots/F1_vs_GWAS_distribution_violin.png", plot = F1vsGWAS_violin, base_asp = 2.5, base_height = 2.5)

library(scales)
show_col(viridis(8, option = "B"))
```
## F0 Plot

"#D44842FF", "#280B54FF", "#9F2A63FF", "#FAC127FF"
```{r}
F0plot <- phe_F0 %>%
  filter(PHE == "FL50" & (YEAR %in% c(2016:2019)) & !(SITE %in% c("OVTN", "MNHT"))) %>%
  left_join(sites, by = "SITE") %>%
  group_by(manu_site, PLANT_ID) %>%
  summarise(MEAS = mean(MEAS, na.rm = TRUE))

F0vsGWAS <- phe_all %>%
  left_join(sites, by = "SITE") %>%
  filter(PHE == "FL50" & WHERE == "GWAS" & !(SITE %in% c("FRMI", "OVTN")) & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = SUBPOP, y = as_date(MEAS, origin = "2019-01-01"))) +
  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_violin(aes(fill = SUBPOP)) + 
  facet_wrap(~manu_site, nrow = 2) +
  geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  scale_fill_manual(values = c("#442C83", "#F47F72")) +
  scale_color_manual(values = c("#D44842FF", "#280B54FF", "#9F2A63FF", "#FAC127FF")) +
  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  xlab("Subpopulation") + ylab("Flowering Date")
F0vsGWAS
save_plot(filename = "../manuscript/plots/F0_vs_GWAS_distribution.png", plot = F0vsGWAS)
```

```{r}
phe_F2 %>%
  group_by(PLANT_ID) %>%
  summarise(count = n()) %>% arrange(count)

FWCRvsGWAS <- phe_all %>%
  filter((WHERE == "FWCR") | (WHERE == "GWAS" & SUBPOP %in% c("Midwest", "Gulf"))) %>%
  filter(PHE == "FL50" & !(SITE %in% c("OVTN", "MNHT", "FRMI"))) %>%
  left_join(sites, by = "SITE") %>%
  dplyr::select(-(Longitude:Note))  %>%
  mutate(POP = case_when(WHERE == "FWCR" & !(PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC", "AxD", "VxW")) ~ "F2",
                         WHERE == "FWCR" & (PLANT_ID %in% c("AP13", "DAC", "VS16", "WBC")) ~ "F0",
                         WHERE == "FWCR" & (PLANT_ID %in% c("AxD", "VxW")) ~ "F1",
                            WHERE == "GWAS" ~ as.character(SUBPOP)))

summary(as_factor(FWCRvsGWAS$POP))
F0plot <- phe_F0 %>%
  filter(PHE == "FL50" & (YEAR %in% c(2016:2019)) & !(SITE %in% c("OVTN", "MNHT"))) %>%
  left_join(sites, by = "SITE") %>%
  group_by(manu_site, PLANT_ID) %>%
  summarise(MEAS = mean(MEAS, na.rm = TRUE))

FWCRvsGWAS %>%
  filter(!(POP %in% c("F0", "F1")) & (YEAR == 2019 | (YEAR == 2018 & SITE == "TMPL"))) %>%
  ggplot(aes(x = POP, y = as_date(MEAS, origin = "2019-01-01"))) +
  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_violin(aes(fill = POP)) + 
  facet_wrap(~manu_site, nrow = 2) +
  geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
  scale_fill_manual(values = c("lightgrey", "#F47F72", "#442C83")) +
  scale_color_manual(values = c("#D44842FF", "grey", "#280B54FF", "#9F2A63FF", "black", "#FAC127FF")) +
  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  xlab("Subpopulation") + ylab("Flowering Date")

FWCRvsGWAS %>%
  filter(!(POP %in% c("F0", "F1")) & (YEAR == 2018 | (WHERE == "GWAS" & YEAR == 2019))) %>%
  ggplot(aes(x = POP, y = as_date(MEAS, origin = "2019-01-01"))) +
  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_violin(aes(fill = POP)) + 
  facet_wrap(~manu_site, nrow = 2) +
  geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
  scale_fill_manual(values = c("lightgrey", "#F47F72", "#442C83")) +
  scale_color_manual(values = c("#D44842FF", "grey", "#280B54FF", "#9F2A63FF", "black", "#FAC127FF")) +
  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  xlab("Subpopulation") + ylab("Flowering Date")

FWCRvsGWAS %>%
  filter(!(POP %in% c("F0", "F1")) & WHERE == "FWCR") %>%
  ggplot(aes(x = as_factor(YEAR), y = as_date(MEAS, origin = "2019-01-01"))) +
  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_violin(aes(fill = as_factor(YEAR))) + 
  facet_wrap(~manu_site, nrow = 2) +
  geom_jitter(data = filter(FWCRvsGWAS, POP %in% c("F0", "F1")), aes(color = PLANT_ID), shape = 4, height = 0) +
  #geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  #geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
  scale_fill_viridis(discrete = TRUE, end = 0.8) + 
  scale_color_manual(values = c("#D44842FF", "grey", "#280B54FF", "#9F2A63FF", "black", "#FAC127FF")) +
  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  xlab("Year") + ylab("Flowering Date") + ylim(c(as_date(50, origin = "2019-01-01"), as_date(300, origin = "2019-01-01")))

save_plot(filename = "../manuscript/plots/FWCR_distribution_by_year.png", plot = last_plot())

FWCRvsGWAS %>%
  filter(!(POP %in% c("F0", "F1")) & WHERE == "FWCR") %>%
  ggplot(aes(x = as_factor(YEAR), y = as_date(MEAS, origin = "2019-01-01"))) +
  geom_hline(yintercept = as_date(173, origin = "2019-01-01"), color = "#FDE725FF", size = 1) +
  geom_jitter(data = filter(FWCRvsGWAS, POP %in% c("F0")), aes(color = PLANT_ID), shape = 4, height = 0) +
  geom_violin(aes(fill = as_factor(YEAR))) + 
  facet_wrap(~manu_site, nrow = 2) +

  #geom_hline(data = F0plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 2) +
  #geom_hline(data = F1plot, aes(yintercept = as_date(MEAS, origin = "2019-01-01"), color = PLANT_ID), linetype = 3) +
  scale_fill_viridis(discrete = TRUE, end = 0.8) + 
  scale_color_manual(values = c("#D44842FF", "#280B54FF", "#9F2A63FF",  "#FAC127FF")) +
  coord_flip() + theme(legend.position = "right", axis.text.x = element_text(hjust = 1, vjust = 1, angle = 45)) +
  xlab("Year") + ylab("Flowering Date") + ylim(c(as_date(50, origin = "2019-01-01"), as_date(300, origin = "2019-01-01")))

save_plot(filename = "../manuscript/plots/FWCR_F0_distribution_by_year.png", plot = last_plot())

```

(How many days behind the parents was each F1 cross, on average, across all 4 years of data?)  
```{r}
F1_F0_diff <- FWCRvsGWAS %>%
  filter((POP %in% c("F0", "F1")) & WHERE == "FWCR") %>%
  group_by(SITE, YEAR, PLANT_ID) %>%
  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>%
  pivot_wider(names_from = PLANT_ID, values_from = meanFL) %>%
  mutate(earlyFL_F1_F0_diff = AxD - DAC,
         lateFL_F1_F0_diff = VxW - VS16,
         earlyFL_F1_F0_diff_g = AP13 - AxD,
         lateFL_F1_F0_diff_g = WBC - VxW
         )  %>%
  ungroup()
F1_F0_diff %>%
  summarise(mean_early_F1_F0_diff = mean(earlyFL_F1_F0_diff, na.rm = TRUE),
            mean_late_F1_F0_diff = mean(lateFL_F1_F0_diff, na.rm = TRUE),
            mean_early_F1_F0_diff_g = mean(earlyFL_F1_F0_diff_g, na.rm = TRUE),
            mean_late_F1_F0_diff_g = mean(lateFL_F1_F0_diff_g, na.rm = TRUE))

sd(F1_F0_diff$earlyFL_F1_F0_diff, na.rm = TRUE)
sd(F1_F0_diff$lateFL_F1_F0_diff, na.rm = TRUE)
sd(F1_F0_diff$earlyFL_F1_F0_diff_g, na.rm = TRUE)
sd(F1_F0_diff$lateFL_F1_F0_diff_g, na.rm = TRUE)
```

```{r}
Gulf_minFL <- FWCRvsGWAS %>%
  filter((POP %in% c("F0"))) %>%
  group_by(SITE, YEAR, PLANT_ID) %>%
  summarise(minFL = min(MEAS, na.rm = TRUE)) %>%
  filter(PLANT_ID %in% c("AP13", "WBC")) %>%
  pivot_wider(names_from = PLANT_ID, values_from = minFL, names_prefix = "minFL_")
Gulf_meanFL <- FWCRvsGWAS %>%
  filter((POP %in% c("F0"))) %>%
  group_by(SITE, YEAR, PLANT_ID) %>%
  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>%
  filter(PLANT_ID %in% c("AP13", "WBC")) %>%
  pivot_wider(names_from = PLANT_ID, values_from = meanFL, names_prefix = "meanFL_")

F2_meanFL <- FWCRvsGWAS %>%
  filter((POP %in% c("F2"))) %>%
  group_by(SITE, YEAR, PLANT_ID) %>%
  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>% 
  left_join(Gulf_meanFL) %>%
  filter(meanFL > meanFL_AP13 | meanFL > meanFL_WBC)
F2_minFL <- FWCRvsGWAS %>%
  filter((POP %in% c("F2"))) %>%
  group_by(SITE, YEAR, PLANT_ID) %>%
  summarise(meanFL = mean(MEAS, na.rm = TRUE)) %>% 
  left_join(Gulf_minFL) %>%
  filter(meanFL >= minFL_AP13 | meanFL >= minFL_WBC)

F2_meanFL %>%
  group_by(YEAR, PLANT_ID) %>%
  tally() %>% arrange(desc(n))
F2_minFL %>%
  group_by( PLANT_ID) %>%
  tally() %>% arrange(desc(n))

478/14672
2308/14672
```
Only 3.26% of F2 flowering dates occurred after the average flowering date for one or both Gulf parents; 13.3% of F2 flowering dates occurred after the minimum flowering date for one or both Gulf parents.

Only 358 of 801 F2 individuals ever flowered after the average flowering date Gulf parent, and only 102 did so in more than one site or year. (13 in three + site/years, 2 in 5-6).

minimum flowering date for Gulf parents
PLANT_ID  n
91	15			
402	13			
105	12			
194	11			
210	11			
248	11			
196	10			
214	10
